---
title: "Group project 2 -- Introduction & EDA"
author: "wifan"
date: "2021/7/12"
output:
  pdf_document: default
  html_document:
    df_print: paged
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, eval = TRUE, warning = FALSE, message = FALSE, comment = NA)
```

```{r libraries}
library(dplyr)
library(maps)
library(RColorBrewer)
library(ggplot2)
library(tidyverse)
library(kableExtra)
library(gridExtra)
library(moderndive)
library(skimr)
library(sjPlot)
library(GGally)
library(knitr)
library(janitor)
library(jtools)
```
# Introduction {#sec:intro}

The original data set about coffee beans has 1134 observations with 8 variables, which comes from the Coffee Quality Database(CQD). These variables are made of the geographical characters of coffee beans, some features of coffee beans, harvested year and their quality class. Each row of data is recorded by batch and it assumes that each batch of tested coffee beans have equal quantity. 

Here, coffee beans quality can be divided into two main categories, 'Good' and 'Poor', according to a chosen cutoff point of quality grade(82.5). In this report, our target is to investigate what extend the features of a batch of coffee affect its quality by using a generalised linear model.

We used R to handle this data set and renamed it as *'data14.csv'*. Since the term 'Qualityclass' is a binary categorical variable, it can be converted to factor format for further analysis and label '1' denotes good quality, label '0' denotes bad quality. The overview of original data is shown below. 
```{r glimpse,echo=FALSE, warning=FALSE}
data14<-read.csv('dataset14.csv')
data14$Qualityclass<-factor(data14$Qualityclass,label = c(1,0))
data14$harvested<-as.factor(data14$harvested)
glimpse(data14)
```

# Exploratory Data Analysis {#sec:EDA}

In order to understand the data, we take a summary of our data set first.
```{r Summary statistics,echo=FALSE, message=FALSE, out.width = '70%',paged.print=TRUE}
#select useful summary statistics of data set
summary.stat<-skim_without_charts(data14)%>%
  select(,-c(5:12,16:18))

#convert into table
options(knitr.kable.NA = '-')
summary.stat%>%
  transmute(type=skim_type,variable=skim_variable,n_missing=n_missing,complete_rate=complete_rate,mean=numeric.mean,sd=numeric.sd,min=numeric.p0,max=numeric.p100)%>%
kable(caption = '\\label{tab:Summary statistics} Summary statistics of the original data set.',
booktabs = TRUE, align = "c",digit=2) %>%
kable_styling(font_size = 10, latex_options = "hold_position")
```
From the information above, we can find that there are some missing values scatter in three columns(1 in column 'country_of_origin', 199 in column 'altitude_mean_meters' and 59 in column 'harvested'). Specifically, the 1 missing value in column 'country_of_origin', which altitude and harvested year are also unknown. There is no enough evidence to determine its origin country. And the missing data is a small amount, so we could remove this row for retaining the data integrity. In addition, the number of missing values in other two columns are relatively large, they will be handled in the next steps. Meanwhile, there are no abnormal values. However, the minimums of aroma grade, flavor grade and acidity grade are all zeros. Although zero is not out of the given range in these columns, these outliers will affect our analysis results and these zeros come from the same batch, so we choose to drop this row of data.

After removing the missing value in column 'country_of_origin' and an outlier, we stored it as *'data14_tidy.csv'*
in R. It still contains 1132 rows and 8 columns and will be used for analyzing in the following part of this report.
```{r tidy data,echo=FALSE}
#tidy data:remove NA in country_of_origin
data14_tidy <-data14 %>% 
  filter(!is.na(country_of_origin))%>%
  filter(aroma!=0)
```

Initially, we had a look at top 5 countries in the number of good quality batches. It shows that the coffee beans origin from Mexico, Colombia and Guatemala have the higher proportion of good quality in general. The number of batches in good quality are 192, 155, and 150 in these three countries respectively.
```{r top5 countries, echo=FALSE, warning=FALSE, out.width = "70%", fig.align = "center", fig.cap = "\\label{fig:bar} Top 5 countries in the number of batches in good quality.", fig.pos = 'H'}
#count the number of batches in good quality by country
coffee_origin<- data14_tidy %>%
  select("country_of_origin", "Qualityclass")%>%
  group_by(country_of_origin)%>%
  summarise(counts = n())
top_5<-coffee_origin%>%
  top_n(5)

#build bar plots 
ggplot(top_5,aes(x=reorder(country_of_origin, -counts), y=counts))+
  geom_bar(stat='identity', fill = 'steelblue')+
  theme(legend.position = "none")+
  geom_text(aes(label=counts,vjust=-0.5))+
  labs(x='Country',y='Count')+
  ylim(0,230)
  
```

Next, we used box plots to explore the relationship between aroma grade, flavor grade, acidity grade and quality class of coffee beans separately.
```{r aroma;flavor;acidity,echo=FALSE, out.width = '70%', fig.align = "center", fig.cap = "\\label{fig:box} The relationship between aroma,flavor,acidity and quality class", fig.pos = 'H'}
p1<-ggplot(data14_tidy,aes(x=Qualityclass,y=aroma,fill=Qualityclass))+
  geom_boxplot()+
  labs(x='Qualityclass',y='Aroma grade')+
  scale_x_discrete(labels=c('Good','Poor'))+
  theme(legend.position = "none")

p2<-ggplot(data14_tidy,aes(x=Qualityclass,y=flavor,fill=Qualityclass))+
  geom_boxplot()+
  labs(x='Qualityclass',y='Flavor grade')+
  scale_x_discrete(labels=c('Good','Poor'))+
  theme(legend.position = "none")

p3<-ggplot(data14_tidy,aes(x=Qualityclass,y=acidity,fill=Qualityclass))+
  geom_boxplot()+
  labs(x='Qualityclass',y='Acidity grade')+
  scale_x_discrete(labels=c('Good','Poor'))+
  theme(legend.position = "none")

grid.arrange(p1,p2,p3,ncol=3)

```
Figure \ref{fig:box} shows that aroma grade, flavor grade and acidity grade of coffee beans in good quality class tend to be higher than that in poor quality class. Meanwhile, there are some outliers in each plot. In terms of this result, these three variables can be considered as influence factors of coffee quality.

In this step, table \ref{tab:category_two_defects} indicates the difference of the number of category 2 type defects in each coffee quality class. The number of category 2 type defects in good quality class(582) is slightly more than that in poor quality class(551).
```{r category_two_defects,echo=FALSE,warning=FALSE}
my_skim <- skim_with(base = sfl(n = length)) 

data14_tidy %>%
group_by(Qualityclass) %>%
select(category_two_defects, Qualityclass) %>%
my_skim() %>%
transmute(Qualityclass=c('Good','Poor'), n=n ) %>%
kable(caption = '\\label{tab:category_two_defects} Summary statistics on the number of category 2 type defects by quality class.',
booktabs = TRUE, linesep = "") %>%
kable_styling(font_size = 10, latex_options = "hold_position")
```

For the purpose of observing whether the harvested year of coffee beans is related to the coffee quality class, a bar plot is made to show the change of the proportion of each quality class during the period from 2010 to 2018. Because the amount of missing data in this column is relatively small, we ignored the missing values and analyzed the data roughly.
```{r harvested year,echo=FALSE, out.width = '70%', fig.align = "center", fig.cap = "\\label{fig:barplot} The proportion of each quality class in every harvested year", fig.pos = 'H'}
data14.year<-data14_tidy %>%
  select(harvested,Qualityclass) %>%
  na.omit()

plot_xtab(data14.year$Qualityclass,data14.year$harvested,show.values=FALSE,show.total=FALSE,
          axis.labels = c('Good','Poor'),legend.title = 'harvested year')
```
As shown in the Figure \ref{fig:barplot}, there are both fluctuation in the proportion of two quality classes during this period. However, we can not find out the regular trends of the two proportions over time from the figure. Noticing the shapes of bar in the figure, it is obvious that the proportions of good quality coffee beans are much higher in 2010, 2011 and 2018 and that in the rest of years are relatively balanced. We speculated that this may be due to the 59 missing values. For this reason, we could infer that harvested year have a little effect on the coffee quality.

Before explore the relationship between the mean altitude of the growers farm and the coffee quality, we need to work with the large amount of missing values in altitude. There are several ways to fill the missing data. The most common two methods are constant replacement and regression interpolation.

First, we checked the correlation between altitude and previous four explanatory variables in figure \ref{fig:corr}. Unfortunately, the correlation coefficients between altitude and other variables are -0.012, -0.01,0.03 and -0.05 respectively, which are very weak relationship, so we will not take regression method into account.
```{r check correlation ,echo=FALSE, out.width = '70%', fig.align = "center", fig.cap = "\\label{fig:corr} Potential factors for altitude", fig.pos = 'H', warning=FALSE, message=FALSE}
corr <- data14_tidy %>%
  select(2:6)%>%
  na.omit()
ggpairs(corr)
```

Then, the constant replacement method will be considered. Due to coffee beans from each country should have similar growing environment and geographical conditions, we choose to use average by each country to deal with missing data of altitude from different area.
```{r deal with missing values,echo=FALSE, warning=FALSE}
#the number of NA,mean and median 'altitude_mean_meters' term group by countries
summary_country_altitude<-data14%>%
  group_by(country_of_origin)%>%
  summarize(na=sum(is.na(altitude_mean_meters)),mean=mean(altitude_mean_meters,na.rm = TRUE))


#build the table
altitude_by_country<-na.omit(summary_country_altitude)%>%
kable(caption = '\\label{tab:Altitude by countries} Summary statistics of the altitudes by country.',
booktabs = TRUE, linesep = "") %>%
kable_styling(font_size = 10, latex_options = "hold_position")


#replace NA in column 'altitude_mean_meters' by mean
countries<-unique(na.omit(summary_country_altitude[summary_country_altitude$na!=0,]$country_of_origin))
for (i in 1:length(countries)){
  data14_tidy[data14_tidy$country_of_origin==countries[i],]$altitude_mean_meters[is.na(data14_tidy[data14_tidy$country_of_origin==countries[i],]$altitude_mean_meters)]<-
    summary_country_altitude$mean[which(summary_country_altitude$country_of_origin==countries[i])]
}

```
After the process of filling the missing data, a box plot is built for visualizing the effect of altitude on coffee qualify class. 
```{r altitude ,echo=FALSE, out.width = '70%', fig.align = "center", fig.cap = "\\label{fig:box2} Altitude by coffee qualify class", fig.pos = 'H', warning=FALSE, message=FALSE}
#for clear visualization, removing 11 points which altitude >6000 in the plot
#nrow(data14_tidy[data14_tidy$altitude_mean_meters>6000,])
ggplot(data14_tidy[data14_tidy$altitude_mean_meters<=6000,],aes(x=Qualityclass,y=altitude_mean_meters,fill=Qualityclass))+
  geom_boxplot()+
  labs(x='Quality class',y='Altitude')+
  scale_x_discrete(labels=c('Good','Poor'))+
  theme(legend.position = "none")
```
Figure \ref{fig:box2} shows that the altitude of the growers farm in good qualify class tend to be slightly higher than that in poor qualify class in general. In addition, the mean altitude of the cultivation of good coffee beans is more concentrated than the distribution of poor qualited coffee beans. Therefore, it can be inferred that the average altitude [Good mean] is more suitable for high-quality output of coffee beans. Manufacturers can choose planting locations to improve the quality of coffee beans.


# Formal Data Analysis {#sec:FDA}

Since we have a binary response variable $Y_i$ we will apply a binary Generalised Linear Model and we will  consider the logit link. From our Exploratory Data Analysis we expect Flavor, Aroma and Acidity to be good predictors of Qualityclass. We also have seen that Altitude and Category Two Defects may have some effect, so they are also included.

We then fit the following model, with Qualityclass = Poor as our baseline category.

$$ln\left(\frac{p_{poor}}{1-p_{poor}}\right) = \alpha + \beta_1 \cdot \textrm{flavor} + \beta_2 \cdot \textrm{aroma} + \beta_3 \cdot \textrm{acidity}+ \beta_4 \cdot \textrm{category two defects} +  \beta_5 \cdot \textrm{altitude}$$


```{r full_model, echo = FALSE, eval = TRUE}
model_full <- glm(Qualityclass ~ flavor + aroma + acidity + category_two_defects + altitude_mean_meters  ,data = data14_tidy,family = binomial(link = "logit"))

summary(model_full)
```
Here we see that the coefficients for flavor, aroma and acidity are all negative, this suggests that an increase in any of these variables results in a decreased chance of being classified as Poor coffee. flavor is suggested to have the greatest impact on decreased chance. Their p-values also suggest significance. Variables category_two_defects and altitude_mean_meters have very small effects and their p-values suggest they are not significant.


We then provide Wald intervals for more inference on the significance of our log(odds).
```{r wald, echo = FALSE, eval = TRUE,fig.align = "center", fig.cap = "\\label{fig:CI} The Wald Confidence Intervals for our log(odds) in Model 1"}
confint.default(model_full) %>% #confint.default calculates wald confidence intervals since logistic regression is asymptotically normal 
  kable()
```

The Wald Intervals for our first 3 variables do not contain 0 so we will keep them in the model. Intervals for the variables category_two_defects and altitude_mean_meters both contain 0, furthering the idea that they are not significant terms. 

We drop them from the model and fit,

$$ln\left(\frac{p_{poor}}{1-p_{poor}}\right) = \alpha + \beta_1 \cdot \textrm{flavor} + \beta_2 \cdot \textrm{aroma} + \beta_3 \cdot \textrm{acidity}$$


```{r score_model, echo = FALSE, eval = TRUE}
model_scores <- glm(Qualityclass ~ flavor + aroma + acidity ,data = data14_tidy,family = binomial(link = "logit"))


summary(model_scores)
```

With confidence intervals of

```{r wald2, echo = FALSE, eval = TRUE,fig.align = "center", fig.cap = "\\label{fig:CI} The Wald Confidence Intervals for our log(odds) in Model 1"}
confint.default(model_scores) %>% #confint.default calculates wald confidence intervals since logistic regression is asymptotically normal 
  kable()
```

And Odds Ratios of

```{r odds_plot2,fig.cap = "\\label{fig:OR} Odds Ratios"}
# plot of odds ratios
plot_model(model_scores, show.values = TRUE, 
title = "", show.p = TRUE)
```

The above figure shows odds very close 0 for each of the 3 variables, that is to say, for each unit increase in score for flavor, aroma and acidity their odds of being classified as poor decrease (by factors close to 0).

```{r histogram,echo=FALSE, out.width = '70%', fig.align = "center", fig.cap = "\\label{fig:hist} The histgrams for aroma, flavor and acidity grades", fig.pos = 'H'}
h1<-ggplot(data14_tidy,aes(x=aroma,fill='steelblue'))+
  geom_histogram(breaks=seq(5,9,0.2))+
  labs(x='Aroma grade')+
  theme(legend.position = "none")
h2<-ggplot(data14_tidy,aes(x=flavor,fill='steelblue'))+
  geom_histogram(breaks=seq(5,9,0.2))+
  labs(x='Flavor grade')+
  theme(legend.position = "none")
h3<-ggplot(data14_tidy,aes(x=acidity,fill='steelblue'))+
  geom_histogram(breaks=seq(5,9,0.2))+
  labs(x='Acidity grade')+
  theme(legend.position = "none")

  grid.arrange(h1,h2,h3,nrow=3)

```


# Conclusion {#sec:Conc}


